Redirecting Resolution Proofs

### 2011-05-10 bib/prooftransform/redirecting-proofs-2/paper.tex
added some missing references

There is a large body of research about making resolution proofs
readable. Earlier work focused on translating detailed resolution
proofs into natural deduction or sequent calculi \cite{xxx, etc.}.
Although they are arguably more readable, these calculi still operate
at the logical level, whereas humans reason mostly at the assertion
level, invoking definitions and lemmas without providing the full
logical details \cite{huang-xxx}. A line of research focused on
transforming natural deduction proofs into assertion-level proofs,
culminating with the systems \textsc{Tramp} \cite{xxx} and Otterfier
\cite{zimmer-et-al-2004}.

@@

As a first step toward more intellible proofs, I looked for a method
to turn contradiction proofs around. It seemed obvious the
method should not be tied to any logic as long as it is classical, or
to any calculus. In particular, it should work on the Isar proofs
generated by Sledgehammer, or directly on first-order TSTP proofs.
Finally, the direct proof should be expressible comfortably in a
block-structured Isar-like language, with case splits and nested
subproofs.

In my PAAR 2010 and IWIL 2010 talks \cite{xxx,xxx}, I sketched a
method for transforming proofs by contradiction into direct proofs.
The method was very simple and simply exploited the contrapositive,
but in the worst case the resulting proof could explode exponentially
in size. In this paper, I propose a new method transform resolution
proofs into direct proofs expressed in a simple Isar-like syntax
(Section~\ref{sec:proof-notations}), based on a simple set of rules
(Sections \ref{sec:introductory-examples} and
\ref{sec:the-algorithm}). The new approach is sound and complete and
has the advantage that each deduction in the ATP proof
gives rise to exactly one deduction in the direct proof. A linear
number of additional steps are introduced in the direct proof, but
these are simple logical rules, such as modus ponens, and do not
require the full power of Metis.
%The approach is illustrated on
%actual ATP proofs found by Sledgehammer (Section~\ref{sec:gallery}).

>>>>

There is a large body of research about making resolution proofs
readable. Earlier work focused on translating detailed resolution
proofs into natural deduction or sequent calculi. % \cite{xxx, etc.}.
Although they are arguably more readable, these calculi still operate
at the logical level, whereas humans reason mostly at the assertion
level, invoking definitions and lemmas without providing the full
logical details. % \cite{huang-xxx}.
A line of research focused on
transforming natural deduction proofs into assertion-level proofs,
culminating with the systems \textsc{Tramp} \cite{meier-2000} and Otterfier
\cite{zimmer-et-al-2004}.

@@

As a first step toward more intelligible proofs, I looked for a method to turn
contradiction proofs around. It seemed obvious the method should not be tied to
any logic as long as it is classical, or to any calculus. In particular, it
should work on the Isar proofs generated by Sledgehammer, or directly on
first-order TSTP proofs. Finally, the direct proof should be expressible
comfortably in a block-structured Isar-like language, with case splits and
nested subproofs.

In my PAAR 2010 and IWIL 2010 talks \cite{blanchette-paulson-2010,paulson-2010},
I sketched a method for transforming proofs by contradiction into direct proofs.
The method was very simple and simply exploited the contrapositive, but in the
worst case the resulting proof could explode exponentially in size. In this
paper, I propose a new method transform resolution proofs into direct proofs
expressed in a simple Isar-like syntax (Section~\ref{sec:proof-notations}),
based on a simple set of rules (Sections \ref{sec:introductory-examples} and
\ref{sec:the-algorithm}). The new approach is sound and complete and has the
advantage that each deduction in the ATP proof gives rise to exactly one
deduction in the direct proof. A linear number of additional steps are
introduced in the direct proof, but these are simple logical rules, such as
modus ponens, and do not require the full power of Metis.
%The approach is illustrated on
%actual ATP proofs found by Sledgehammer (Section~\ref{sec:gallery}).
